---
title: "Geography 176A"
author: "[Jordan Molinski](https://github.com/J-Molinski)"
subtitle: 'Lab 03: Distances and the Border Zone'
output:
  html_document:
    theme: journal
---

```{r}

# SPDS
library(tidyverse)
library(sf)
library(units)

# Data
library(USAboundaries)
library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)
library(kableExtra)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, out.width = "75%", fig.align = 'center')
```

### Question 1 : Creating Manipulatable Datasets and applying Coordinate Reference Systems
```{r}
eqdc = "+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
```

```{r}
conus = USAboundaries::us_states(resolution = "low") %>%
  filter(!state_name %in% c("Alaska", "Puerto Rico", "Hawaii")) %>%
  st_transform(eqdc)
```

```{r}
worldboundaries = rnaturalearth::countries110 %>%
  st_as_sf(worldboundaries) %>%
  filter(admin %in% c("Canada", "Mexico", "United States of America")) %>%
  st_transform(eqdc)
```

```{r}
uscities = readr::read_csv("../data/uscities.csv") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  filter(!state_name %in% c("Alaska", "Puerto Rico", "Hawaii")) %>%
  st_transform(eqdc)
```


### Question 2: Calculating Distances

##2.1 City to US Border
```{r}
conusunion = st_union(conus) %>%
  st_cast("MULTILINESTRING")
```

```{r}
uscities1 = uscities %>%
  mutate(dist_to_usBorder = st_distance(uscities, conusunion)) %>%
  select(state_name, city, dist_to_usBorder) %>%
  slice_max(dist_to_usBorder, n=5)
 
city_to_border = knitr::kable(uscities1,
              caption = "The 5 US Cities Furthest from a National Border or Coastline",
              col.names = c("State", "City", "Distance to National Border or Coastline (m)", "Geometry"))

city_to_border
```


##2.2 City to State Border
```{r}
conuscombine = st_combine(conus) %>%
  st_cast("MULTILINESTRING")
```

```{r}
city_to_stateborder = uscities %>%
  mutate(dist_to_stateborder = st_distance(uscities, conuscombine)) %>%
  select(state_name, city, dist_to_stateborder) %>%
  slice_max(dist_to_stateborder, n=5)

city_to_state = knitr::kable(city_to_stateborder,
                             caption = "The 5 US Cities Furthest from a State Border",
                             col.names = c("State", "City", "Distance to State Border (m)", "Geometry"))

city_to_state
```

##2.3 City to Mexican Border
```{r}
mexico = worldboundaries %>%
  filter(admin == "Mexico")
  
mexico_combine = st_combine(mexico) %>%
  st_cast("MULTILINESTRING")
```


```{r}
city_to_mexico = uscities %>%
  mutate(dist_to_mexico = st_distance(uscities, mexico_combine)) %>%
  select(state_name, city, dist_to_mexico) %>%
  slice_max(dist_to_mexico, n=5)

city_to_mexicotable = knitr::kable(city_to_mexico,
                                   caption = "The 5 US Cities Furthest from Mexico",
                                   col.names = c("State", "City", "Distance to Mexico (m)", "Geometry"))

city_to_mexicotable
```

##2.4 City to Canadian Border
```{r}
canada = worldboundaries %>%
  filter(admin =="Canada")

canada_combine = st_combine(canada) %>%
  st_cast("MULTILINESTRING")
```

```{r}
city_to_canada = uscities %>%
  mutate(dist_to_canada = st_distance(uscities, canada_combine)) %>%
  select(state_name, city, dist_to_canada) %>%
  slice_max(dist_to_canada, n=5)

city_to_canadatable = knitr::kable(city_to_canada,
                              caption = "The 5 US Cities Furthest from Canada",
                              col.names = c("State", "City", "Distance to Mexico (m)", "Geometry"))

city_to_canadatable
```


### Question #3 Visualizing Distance Data

##3.1 Data
```{r}
bigcities = uscities %>%
  select(state_name, city, population) %>%
  st_transform(eqdc) %>%
  slice_max(population, n = 10)
```


```{r}
plot1 = ggplot() +
  geom_sf(data = conuscombine) +
  geom_sf(data = mexico_combine) +
  geom_sf(data = canada_combine) +
  geom_sf(data = bigcities, col = "red") +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(
    data = bigcities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
  ) 
ggsave(plot1, filename = "../img/Q3-Plot1.png")
plot1

```


## 3.2 -- City to US border map
```{r}
uscities2 = uscities %>%
  mutate(dist_to_usBorder = st_distance(uscities, conusunion)) %>%
  select(state_name, city, dist_to_usBorder) %>%
  drop_units()
```

```{r}
plot2 = ggplot() +
  geom_sf(data = conusunion) +
  geom_sf(data = uscities2, aes(col = dist_to_usBorder), size = .1) +
  geom_sf(data = uscities1, col = "blue") + 
  scale_color_gradient(low = "gray", high = "red") +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(
    data = uscities1,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates"
  )

ggsave(plot2, filename = "../img/Q3-Plot2.png")
plot2
  
```

## 3.3 -- City to State Border Map

```{r}
city_to_stateborder1 = uscities %>%
  mutate(dist_to_stateborder = st_distance(uscities, conuscombine)) %>%
  select(state_name, city, dist_to_stateborder) %>%
  drop_units()
```

```{r}
plot3 = ggplot() +
  geom_sf(data = conuscombine) +
  geom_sf(data = city_to_stateborder1, aes(col = dist_to_stateborder), size = .1) +
  geom_sf(data = city_to_stateborder, col = "blue") + 
  scale_color_gradient(low = "gray", high = "red") +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(
    data = city_to_stateborder,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates"
  )

ggsave(plot3, filename = "../img/Q3-Plot3.png")
plot3
```

## 3.4 -- Equidistant to Canada and Mexico
```{r}
mex_can = uscities %>%
  mutate(dist_to_canada = st_distance(., canada_combine), 
         dist_to_canada = set_units(dist_to_canada, "km"), 
         dist_to_canada = drop_units(dist_to_canada)) %>%
  mutate(dist_to_mexico = st_distance(., mexico_combine),
         dist_to_mexico = set_units(dist_to_mexico, "km"),
         dist_to_mexico = drop_units(dist_to_mexico)) %>%
  mutate(equidist = abs(dist_to_mexico - dist_to_canada)) %>%
  select(city, state_name, population, equidist) 

mex_can5 = mex_can %>%
  filter(equidist <= 100) %>%
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = mex_can, aes(col = equidist), size = .1) +
  geom_sf(data = conusunion) +
  scale_color_gradient(low = "gray", high = "red") +
  gghighlight::gghighlight(equidist <= 100) +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(
    data = mex_can5,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates"
  )
 
```
filter()
group_by(state)
slice()
ungroup()


### Question #4 == Application

## 4.1 -- Table of information
```{r}
worldboundariescombine = st_combine(worldboundaries) %>%
  st_cast("MULTILINESTRING")

cities100 = uscities %>%
  mutate(dist_to_bord = st_distance(., worldboundariescombine), 
         dist_to_bord = set_units(dist_to_bord, "km"), 
         dist_to_bord = drop_units(dist_to_bord)) 

cities100pop = cities100 %>%
  st_drop_geometry() %>%
  summarise(population) %>%
  sum()

citiespercent = cities100pop/328200000
tablesub = data.frame(num_cities = 7481,
                      pop = cities100pop,
                      per = citiespercent,
                      match = FALSE)
kable(tablesub,
             col.names = c("Number of Cities", "Total Population", "Percent", "Match to ACLU"),
             caption = "Comparison to ACLU") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

  
```

## 4.2 -- Mapping Border Zone
```{r}
topcities = cities100 %>%
  filter(dist_to_bord <= 100) %>%
  group_by(state_name) %>%
  slice_max(population, n = 1) %>% 
  ungroup()
```

```{r}
ggplot() +
  geom_sf(data = cities100, aes(col = dist_to_bord), size = .3) +
  geom_sf(data = worldboundariescombine) +
  geom_sf(data = conuscombine) +
  scale_color_gradient(low = "orange", high = "darkred") +
  gghighlight::gghighlight(dist_to_bord <= 100) +
  ggthemes::theme_map() +
  ggrepel::geom_label_repel(
    data = topcities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates"
  )

```


